Reproductive Allocation Calculations

### 2015-03-25 leaflifespan.Rmd
-script runs all the way and generate html file -script generates 3 dataframes with summaries by individual, species*age, species; each including all applicable data - strange problems: still one use of tagID in input file, and reruning all investment calculations if I close and re-open R - need to figure out how to avoid duplicating columns when merging data (in a nicer way than I've implemented)

---
output: html_document
---
#file names
- leaf_counts is file with calculated leaf counts and leaf lifespan data
- LLSummary is file with means of LL - and other variables - by species and age; LMA data is merged to this file
- subsets of LLSummary are created to omit babies or HATE for various analyses

# general results

@@

@@

```{r, echo=FALSE,results='hide', message=FALSE}
# load packages

library(remake)
create_bindings()

@@

# leaf lifespan data
leaf_lifespan <- filter(leafLifespan_raw, dont_use!="dead" & dont_use!="dont_use")
              select(species,age,site,tag_ID,d_start,d_end,length_start,new_length,segment,lvs_start_count,lvs_end_count,lvs_new_count,new_but_shed_count)

# where counts did not previously exists, set as 0; CHECK WITH DANIEL THAT "V" IS A SET "NAME" IN THESE FUNCTIONS
for(v in c("lvs_start_count","lvs_end_count","lvs_new_count","new_but_shed_count")) {
  i <- is.na(leaf_lifespan[[v]])
  leaf_lifespan[[v]][i] <- 0
}

@@

# get spp average for all individuals
raw_leaf_lifespan$count_per_length <- leaves_per_length_spp$count_per_length[match(raw_leaf_lifespan$species, leaves_per_length_spp$species)]

# if possible use data for that individual, if it exists; this code substitutes in species count per length #s for individuals where this data is available
i <- !is.na(raw_leaf_lifespan$mm_lvs)
raw_leaf_lifespan$count_per_length[i] <- (raw_leaf_lifespan$count_lvs_spec/raw_leaf_lifespan$mm_lvs)[i]

@@

#get total leaf counts
for(v in c("lvs_start_length","lvs_start_count","lvs_end_length","lvs_end_count","lvs_new_length","lvs_new_count","new_but_shed_count")) {
  i <- is.na(raw_leaf_lifespan[[v]])
  raw_leaf_lifespan[[v]][i] <- 0
  }
raw_leaf_lifespan$lvs_start <- raw_leaf_lifespan$lvs_start_count + (raw_leaf_lifespan$lvs_start_length * raw_leaf_lifespan$count_per_length)
raw_leaf_lifespan$lvs_end <- raw_leaf_lifespan$lvs_end_count + (raw_leaf_lifespan$lvs_end_length * raw_leaf_lifespan$count_per_length)
raw_leaf_lifespan$lvs_new <- raw_leaf_lifespan$lvs_new_count + (raw_leaf_lifespan$lvs_new_length * raw_leaf_lifespan$count_per_length)

#calculating leaf lifespan
raw_leaf_lifespan$LL_bd <- (raw_leaf_lifespan$lvs_start)/(((raw_leaf_lifespan$lvs_start)-(raw_leaf_lifespan$lvs_end))+(raw_leaf_lifespan$lvs_new)/2)
raw_leaf_lifespan$LL_death <- (raw_leaf_lifespan$lvs_start)/((raw_leaf_lifespan$lvs_start)-(raw_leaf_lifespan$lvs_end))
raw_leaf_lifespan$LL_birth <- (raw_leaf_lifespan$lvs_start)/(raw_leaf_lifespan$lvs_new)

raw_leaf_lifespan$LL_death[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA
raw_leaf_lifespan$LL_bd[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA
raw_leaf_lifespan$LL_birth[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA

#write.csv(raw_leaf_lifespan, file = "raw_leaf_lifespan.csv")
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
#adding leaf, stem weight data from InvestmentSummary to raw_leaf_lifespan dataframe; can't RUN LINES 95-108 UNTIL INVESTMENT SUMMARY OPEN - THEN PROBLEM WITH PLYR
#raw_leaf_lifespan <- left_join(raw_leaf_lifespan, InvestmentSummary, by = "individual")

#adding columns to raw_leaf_lifespan to look at change across year
raw_leaf_lifespan$change_shoot_diam <- raw_leaf_lifespan$d_end - raw_leaf_lifespan$d_start
#raw_leaf_lifespan$change_basal_diam <- raw_leaf_lifespan$FinalBasalDiam - raw_leaf_lifespan$StartBasalDiam
raw_leaf_lifespan$change_shoot_area <- (3.14*((raw_leaf_lifespan$d_end/2)^2)) - (3.14*((raw_leaf_lifespan$d_start/2)^2))
#raw_leaf_lifespan$change_basal_area <- (3.14*((raw_leaf_lifespan$FinalBasalDiam/2)^2)) - (3.14*((raw_leaf_lifespan$StartBasalDiam/2)^2))

#removng crazy numbers; probably better to do this for each analysis than overall, because removing indiviiduals from all three at once
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_bd>0)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_bd<6)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_birth<8)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_death<8)

#creating summary files
#leaf lifespan summary by species, age
LLSummary <- raw_leaf_lifespan %>%
  filter(LL_bd>0 & LL_bd<6 & LL_birth<8 & LL_death<8) %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth,new_length, lvs_end, lvs_end_total)

@@

#leaf lifespan summary by species
LL_spp <- raw_leaf_lifespan %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth)

@@

#merge LMA, wood density, seed size data into LLSummary dataframe
LLSummary <- merge(LLSummary, LMA_summary, by.x=c("species","age"), by.y=c("species","age"))
LLSummary <- merge(LLSummary, Wood_summary, by.x=c("species"), by.y=c("species"))
LLSummary <- merge(LLSummary, seedsize, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, LMA_spp, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, Wood_summary, by.x=c("species"), by.y=c("species"))
LL_spp <- merge(LL_spp, seedsize, by.x=c("species"), by.y=c("species"))
#still to add, max H from Investment data, absolute and relative growth measures

#plotting and stats
#creating lists of information for plotting symbols, colors, labels
pch.spp <- c(0,1,2,15,16,17,18,0,1,2,15,16,17,18)
col.spp <- c("red", "dark green", "blue", "purple", "orange", "light green","salmon","turquoise","grey", "black","light blue","brown","yellow","salmon")
col.age <-c("purple","pink","green","orange","blue","dark green")
pch.age <- c(1,1,16,16,16,16)
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")
labels.age <- c("1.4","2.4","5","7","9","32")

#LL vs LMA
#remove HATE, babies
#instead of separate dataframes, can probably incorporate this information directly into plotting and analysis commands
#LL_nobabies <- subset(raw_leaf_lifespan, age>2)
#LMA_nobabies <- subset(LMA, age>2)
#LL_noHate <- subset(LLSummary, species != "HATE")
#LL_noHatebabies <- subset(LL_nobabies, species != "HATE")
```

# relationship between LMA, LL by age, species

For 7 year old plants
```{r, echo=FALSE}
plot(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7), xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA", las=1, pch=16)
text(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7), labels=species, pos=1, offset=-0.9, cex=0.9,col="blue")
text(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7 & species=="HATE"), labels=species, pos=1, offset=-0.9, cex=0.9,col="red")
mod <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,species!="HATE" & age==7))
summary(mod)
abline(mod)
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_basal_diam ~ leaf_new_total, raw_leaf_lifespan, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in basal diameter")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(leaf_new_total) ~ log(change_basal_diam+1) + species +age, raw_leaf_lifespan)
summary(mod)

plot(change_shoot_diam ~ leaf_new_total, raw_leaf_lifespan, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in diameter of leader")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(leaf_new_total) ~ log(change_shoot_diam+10) + species +age, data= subset(raw_leaf_lifespan, !is.na(change_shoot_diam)))
summary(mod)
```

#LL vs LMA using species by site means

Overall plot shows a lack of relationship when HATE included. This is because it has such a high LMA. The youngest and oldest plants have no pattern, presumably because LL data are inaccurate

@@

col.LMA <-c("black","purple", "pink", "light green", "orange","blue", "light blue","red")
labels.LMA <-c("all except HATE","age 1.4", "age 2.4", "age 5", "age 7", "age 9", "age 32", "including HATE")
plot(LL_birth_mean ~ LMA_mean, LLSummary, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(LLSummary, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(LLSummary, species=="HATE"), pch=17)
#need log transformation to get these to plot onto axes
mod <- lm(LL_birth_mean ~ LMA_mean, LLSummary)
abline(mod, col="red", lty=2)
mod_no_HATE <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary, species!="HATE"))
abline(mod_no_HATE, lwd=2)
mod_no_HATE5 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==5&species!="HATE"))
abline(mod_no_HATE5, col="light green", lwd=1.5)
mod_no_HATE7 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7&species!="HATE"))
abline(mod_no_HATE7, col="orange", lwd=1.5)
mod_no_HATE9 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==9&species!="HATE"))
abline(mod_no_HATE9, col="blue", lwd=1.5)
mod_no_HATE2.4 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==2.4&species!="HATE"))
abline(mod_no_HATE2.4, col="pink", lwd=1.5)
mod_no_HATE1 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age<2.4&species!="HATE"))
abline(mod_no_HATE1, col="purple", lwd=1.5)
mod_no_HATE32 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==32&species!="HATE"))
abline(mod_no_HATE32, col="light blue", lwd=1.5)

legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1.4)
summary(mod)
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
mod <- lm(LL_birth ~ age + species +age*species, data=subset(raw_leaf_lifespan,age>3 & LL_birth!="Inf"))
summary(mod)

#WHY ARE THE FOLLOWING TWO LINES OF CODE NEEDED?
raw_leaf_lifespan %>% group_by(species) %>%
    summarise(mean(LL_birth))

#where does the [spp] comes from? This is the first place in the script it is used like this - I assume it could be any name?
plot_LLvAge <- function(spp) {
  plot(LL_birth ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leaf lifespan (leaf births)", cex.axis=.8, las=1,
     main=labels.spp[spp], xlim=c(1,35), ylim=c(0,8), log="x", pch=16)
  points(LL_birth_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((LL_birth_mean+LL_birth_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((LL_birth_mean-LL_birth_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

@@

```{r, echo=FALSE, message=FALSE}
#plot(LMA_mean ~ as.factor(species), data=subset(LLSummary,age>3), xlab="species", ylab="LMA", cex.axis=.8, las=1)
#mod <- lm(LMA ~ age +species + age*species, LMA)
#summary(mod)

@@

plot_LMAvAge <- function(spp) {
  plot(LMA ~ age, data=subset(LMA,species==spp), xlab="age", ylab="LMA", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp])
  points(LMA_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((LMA_mean+LMA_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((LMA_mean-LMA_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
  #I want to add a line of code that plots over the means for each age. I did this above by calling in another file, but wondering about on-the-spot calculations
}

@@

for(n in names(labels.spp)) {
  plot_LMAvAge(n)
  mod_LMAvAge(n)
  summary(mod_LMAvAge(n))
}

@@

plot_lfcountvAge <- function(spp) {
  plot(lvs_end_total ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leaf count along shoot", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp])
  points(lvs_end_total_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((lvs_end_total_mean+lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((lvs_end_total_mean-lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

mod_lfcountvAge <- function(spp) {
  lm(lvs_end ~ age, data=subset(raw_leaf_lifespan,age>2.4 & species==spp))
}

for(n in names(labels.spp)) {
  plot_lfcountvAge(n)
  mod_lfcountvAge(n)
  summary(mod_lfcountvAge(n))
}

>>>>

---
output: html_document
---
#file names
- leaf_counts is file with calculated leaf counts and leaf lifespan data
- SummarySppAge is file with means of LL - and other variables - by species and age; LMA data is merged to this file
- subsets of SummarySppAge are created to omit babies or HATE for various analyses

#variables in my dataset
Measures of biomass and growth include:
- basal diameter
- "leader" diameter
- total biomass
- leaf mass (calculations not yet done)
- leaf area (calculations not yet done)
- stem biomass (calculations not yet done)
- leaf count along leaders
- shoot extension

* all these can be plotted against age as actual values or as change from a year earlier
* these measures are all done at the level of "individual" with 7 individuals of each species at each of 6 sites

Reproduction measures include:
- total reproductive investment (biomass)
- RA (ratio of reproductive investment to total surplus energy)
- investment in each of 5 different categories of reproductive investment:
  * tissues aborted pre-pollination (ie. aborted bud)
  * tissues that reach their final pre-pollination stage (i.e. open flower)
  * tissues aborted after pollination (i.e. aborted immature fruit)
  * tissues associated with packaging and dispersal of successful, mature fruit (i.e. cones, wings, seed pods)
  * the propagule itself (either the seed or the seed with fused fruit tissue)
- % investment in accessory tissues versus the propagule
- seed set (calculations not yet done)

  !!!TO DISCUSS: I only have my own seed weights for species where the seed isn't fused to any fruit tissue. However, I do have seed weights from the literature and global databases for all species and calculations could be redone based on actual seed weights for all species.!!!

Traits are:
- leaf lifespan (individuals)
- LMA (species * age = site)
- wood density (species)
- maximum height (species or site)
- seed size (species)

!!!TO DISCUSS: There are many way to calculate leaf lifespan, resulting in somewhat different patterns!!!

#previously observed patterns that are - fortunately - also true for my dataset
- correlation between LMA and leaf lifespan. (True when HATE is removed.)
- larger seeds have higher accessory costs

##GENERAL THOUGHTS (and conclusions)

#Leaf lifespan, LMA, leaf area, leaf weight changes with age:

In the field, it seemed that older plants had fewer leaves on their shoots, indicating either decreasing lower leaf lifespan (opposite accepted trend) and/or decreasing shoot extension with age.

* results show no overall change in leaf lifespan with age, although some species have slightly increasing leaf lifespan
* both the number of leaves on a shoot and shoot extension decline with age, with some species having quite dramatic shifts (and others no shift)
* TO DO: Total plant leaf area (and leaf number) with age, scaled to basal area and to stem weight
* TO DO: Plot leaf weight against height (same plot as Daniel used in the data paper in Ecology)

Revised thoughts: The leading shoots show a slight decline in leaf number with age. Overall, the plant will look a lot "leggier" because non-leading shoots with show less new growth and even fewer leaves. This pattern should be captured in by looking at shifts in the ratio of leaf count (or area) to basal stem area (ir total stem weight) across ages.

# overall LL, LMA results

@@

#Accessory costs

Accessory costs account for the vast majority of my plant's reproductive investment - and presumably for other perennial plants worldwide - but this is an often ignored pool.
* Overall, larger seeds have higher accessory costs (as has been shown for other angiosperm species and given as one of the reasons why angiosperms can have smaller seeds than gymnosperms)

While overall accessory costs are not clearly affected by age, several of the sub-categories do change with age:
- proportion of reproductive energy investment in aborted buds and flowers declines after the first few reproductive years

#Correlation between each of the five "traits" and growth and reproductive measures
* TO DO: Re

#discuss with Daniel
- once we are certain about leaf lifespan calculations, those calculations should go into a separate file, since that is background processing, but like calculating investment. Then this becomes the file where variables are summarized, plotted analyzed.
- make sure regressions exist (or are done), to separately look at leaf and stem biomass
- seed weight vs. propagule weight dilemma
- sites that are aged 1.35 vs 1.4 years

@@

# load packages and "open" dataframes to use
library(remake)
create_bindings()

@@

harvest <- filter(HarvestData, use_status=="use" & segment==1 & start_end=="end")

investment <- ReproductionAllocation_all
names(investment) <- c("individual","ReproInv","age","GrowthInv","StartWeight","FinalWeight","StartBasalDiamAv","FinalBasalDiamAv","TotalInvestment","RA","species")
#still have tagId as label
#remaking this from scratch

accessory <- AccessoryCosts_all

leafLifespan <- filter(leafLifespan_raw, dont_use!="dead" & dont_use!="dont_use")
leafLifespan$individual <- as.factor(leafLifespan$individual)
#loading this column as character, so using this for now

# Calculating Leaf Lifespan

# where counts did not previously exists, set as 0; CHECK WITH DANIEL THAT "V" IS A SET "NAME" IN THESE FUNCTIONS

leaf_lifespan_temp <- leafLifespan %>%
              select(species,age,site,individual,d_start,d_end,length_start,new_length,segment,lvs_start_count,lvs_end_count,lvs_new_count,new_but_shed_count)

for(v in c("lvs_start_count","lvs_end_count","lvs_new_count","new_but_shed_count")) {
  i <- is.na(leaf_lifespan_temp[[v]])
  leaf_lifespan_temp[[v]][i] <- 0
}

@@

# get spp average for all individuals
leafLifespan$count_per_length <- leaves_per_length_spp$count_per_length[match(leafLifespan$species, leaves_per_length_spp$species)]

# if possible use data for that individual, if it exists; this code substitutes in species count per length #s for individuals where this data is available
i <- !is.na(leafLifespan$mm_lvs)
leafLifespan$count_per_length[i] <- (leafLifespan$count_lvs_spec/leafLifespan$mm_lvs)[i]

@@

#get total leaf counts
for(v in c("lvs_start_length","lvs_start_count","lvs_end_length","lvs_end_count","lvs_new_length","lvs_new_count","new_but_shed_count")) {
  i <- is.na(leafLifespan[[v]])
  leafLifespan[[v]][i] <- 0
  }
leafLifespan$lvs_start <- leafLifespan$lvs_start_count + (leafLifespan$lvs_start_length * leafLifespan$count_per_length)
leafLifespan$lvs_end <- leafLifespan$lvs_end_count + (leafLifespan$lvs_end_length * leafLifespan$count_per_length)
leafLifespan$lvs_new <- leafLifespan$lvs_new_count + (leafLifespan$lvs_new_length * leafLifespan$count_per_length)

#calculating leaf lifespan
leafLifespan$LL_bd <- (leafLifespan$lvs_start)/(((leafLifespan$lvs_start)-(leafLifespan$lvs_end))+(leafLifespan$lvs_new)/2)
leafLifespan$LL_death <- (leafLifespan$lvs_start)/((leafLifespan$lvs_start)-(leafLifespan$lvs_end))
leafLifespan$LL_birth <- (leafLifespan$lvs_start)/(leafLifespan$lvs_new)

leafLifespan$LL_death[which(is.infinite(leafLifespan$LL_death))] <- NA
leafLifespan$LL_bd[which(is.infinite(leafLifespan$LL_death))] <- NA
leafLifespan$LL_birth[which(is.infinite(leafLifespan$LL_death))] <- NA

#adding investment and accessory costs data to leafLifespan dataframe to create a dataframe with all individual level data
#SummaryInd <- left_join(leafLifespan, investment, by = "individual")
#SummaryInd <- left_join(SummaryInd, accessory, by = "individual")
#SummaryInd <- left_join(SummaryInd, harvest, by = "individual")
#not sure why the above text isn't working... using merge instead

SummaryInd <- merge(leafLifespan, investment, by.x=c("individual"), by.y=c("individual"))
SummaryInd <- subset(SummaryInd, select = -c(age.y,species.y) )
SummaryInd <- merge(SummaryInd, accessory, by.x=c("individual"), by.y=c("individual"))
SummaryInd <- subset(SummaryInd, select = -c(age,species) )
SummaryInd <- merge(SummaryInd, harvest, by.x=c("individual"), by.y=c("individual"))
#problem here, because duplicates number of entries
SummaryInd <- subset(SummaryInd, select = -c(age.x,species.x) )

#adding variables to look at change across years
SummaryInd$change_shoot_diam <- SummaryInd$d_end - SummaryInd$d_start
SummaryInd$change_basal_diam <- SummaryInd$FinalBasalDiam - SummaryInd$StartBasalDiam
SummaryInd$change_shoot_area <- (3.14*((SummaryInd$d_end/2)^2)) - (3.14*((SummaryInd$d_start/2)^2))
SummaryInd$change_basal_area <- (3.14*((SummaryInd$FinalBasalDiam/2)^2)) - (3.14*((SummaryInd$StartBasalDiam/2)^2))
SummaryInd$lvs_end_total <- SummaryInd$lvs_end + SummaryInd$lvs_new
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
#summarizing data by species, age

#leaf lifespan summary by species, age
LL_summary <- SummaryInd %>%
  filter(LL_bd>0 & LL_bd<6 & LL_birth<8 & LL_death<8) %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth,new_length, lvs_end, lvs_end_total, change_shoot_diam, change_basal_diam, change_shoot_area, change_basal_area)

#accessory tissue summary by species, age
accessory_summary <- accessory %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), PrePol_A, PrePol_S, PostPol_A, PD, Prop)

#investment summary by species, age
investment_summary <- investment %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), GrowthInv, FinalWeight, FinalBasalDiamAv, TotalInvestment, RA)

#harvest summary by species, age
harvest_summary <- harvest %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), total_stem_weight, total_leaf_weight, height)

#RA for each species/age combination                 
maxRA <- investment %>%
  group_by(species, age) %>%
  summarise_each(funs(max), RA)
names(maxRA)<-c("species","age","maxRA")

#maximum H for each species/age combination                 
maxH <- harvest %>%
  group_by(species, age) %>%
  summarise_each(funs(max), height)
names(maxH)<-c("species","age","maxH")

@@

#leaf lifespan summary by species
LL_spp <- leafLifespan %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth)

@@

#maximum height for each species
MaxH_spp <- harvest %>%
  group_by(species) %>%
  summarise_each(funs(max), height)
names(LMA_spp)<-c("species", "maxH_spp")

SummaryInd <- merge(SummaryInd,Wood_summary, by.x=c("species"), by.y=c("species"))
SummaryInd <- merge(SummaryInd,LMA_summary, by.x=c("species","age"), by.y=c("species","age"))
SummaryInd <- merge(SummaryInd,MaxH_spp, by.x=c("species"), by.y=c("species"))
SummaryInd <- merge(SummaryInd,seedsize, by.x=c("species"), by.y=c("species"))
#TO DO: REMOVE DUPLICATE COLUMNS FOR AGE and SPECIES

#merge various data summaries into SummarySppAge dataframe
SummarySppAge <- merge(LL_summary, LMA_summary, by.x=c("species","age"), by.y=c("species","age"))
SummarySppAge <- merge(SummarySppAge, Wood_summary, by.x=c("species"), by.y=c("species"))
SummarySppAge <- merge(SummarySppAge, seedsize, by.x=c("species"), by.y=c("species"))
SummarySppAge <- merge(SummarySppAge, MaxH_spp, by.x=c("species"), by.y=c("species"))
SummarySppAge <- merge(SummarySppAge, maxH, by.x=c("species","age"), by.y=c("species","age"))
SummarySppAge <- merge(SummarySppAge, investment_summary, by.x=c("species","age"), by.y=c("species","age"))
SummarySppAge <- merge(SummarySppAge, accessory_summary, by.x=c("species","age"), by.y=c("species","age"))
SummarySppAge <- merge(SummarySppAge, harvest_summary, by.x=c("species","age"), by.y=c("species","age"))
#TO DO: REMOVE DUPLICATE COLUMNS FOR AGE

#merge various data summaries into SummarySpp dataframe
SummarySpp <- merge(LL_spp, LMA_spp, by.x=c("species"), by.y=c("species"))
SummarySpp <- merge(SummarySpp, Wood_summary, by.x=c("species"), by.y=c("species"))
SummarySpp <- merge(SummarySpp, seedsize, by.x=c("species"), by.y=c("species"))
SummarySpp <- merge(SummarySpp, maxH, by.x=c("species"), by.y=c("species"))

#WONDERING IF I SHOULD AT THIS POINT HAVE THE DATA FRAME CALLED "sUMMARYINDIVIDUAL" INSTEAD, SINCE NOW CONTAINS FAR MORE THAN LL DATA

#plotting and stats
#creating lists of information for plotting symbols, colors, labels
pch.spp <- c(0,1,2,15,16,17,18,0,1,2,15,16,17,18)
col.spp <- c("red", "dark green", "blue", "purple", "orange", "light green","salmon","turquoise","grey", "black","light blue","brown","yellow","salmon")
col.age <-c("purple","pink","green","orange","blue","dark green")
pch.age <- c(1,1,16,16,16,16)
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")
labels.age <- c("1.4","2.4","5","7","9","32")
names(labels.age) <- c("1.4","2.4","5","7","9","32")
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_basal_diam ~ lvs_end_total, SummaryInd, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in basal diameter")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(lvs_end_total) ~ log(change_basal_diam+1) + species +age, SummaryInd)
summary(mod)

plot(change_shoot_diam ~ lvs_end_total, SummaryInd, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in diameter of leader")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(lvs_end_total) ~ log(change_shoot_diam+10) + species +age, data= subset(SummaryInd, !is.na(change_shoot_diam)))
summary(mod)
```

#LL vs LMA using species by site means
Overall plot shows a lack of relationship when HATE included. This is because it has such a high LMA. The youngest and oldest plants have no pattern, presumably because LL data are inaccurate

@@

col.LMA <-c("black","purple", "pink", "light green", "orange","blue", "light blue","red")
labels.LMA <-c("all except HATE","age 1.4", "age 2.4", "age 5", "age 7", "age 9", "age 32", "including HATE")
plot(LL_birth_mean ~ LMA_mean, SummarySppAge, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(SummarySppAge, species=="HATE"), pch=17)
#need log transformation to get these to plot onto axes
mod <- lm(LL_birth_mean ~ LMA_mean, SummarySppAge)
abline(mod, col="red", lty=2)
mod_no_HATE <- lm(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE"))
abline(mod_no_HATE, lwd=2)
legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1.4)
summary(mod_no_HATE)

#plot_by_age <- function(age) {
#  mod <- lm(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge,age==age&species!="HATE")
#  abline(mod, col=col.age[age], lwd=1.5)
#}

#for(n in names(labels.age)) {
#  plot_by_age(n)
#}

# Relationship between LMA, LL for 7 year old plants
# There would be a similar plot for any other age as well. As with the overall LMA analyses, HATE needs to be removed.

plot(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge,age==7), xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA", las=1, pch=16)
text(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge,age==7), labels=species, pos=1, offset=-0.9, cex=0.9,col="blue")
text(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge,age==7 & species=="HATE"), labels=species, pos=1, offset=-0.9, cex=0.9,col="red")
mod <- lm(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge,species!="HATE" & age==7))
summary(mod)
abline(mod)
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
mod <- lm(LL_birth ~ age + species +age*species, data=subset(SummaryInd,age>3 & LL_birth!="Inf"))
summary(mod)

#WHY ARE THE FOLLOWING TWO LINES OF CODE NEEDED?
leafLifespan %>% group_by(species) %>%
    summarise(mean(LL_birth))

#where does the [spp] comes from? This is the first place in the script it is used like this - I assume it could be any name?
plot_LLvAge <- function(spp) {
  plot(LL_birth ~ age, data=subset(SummaryInd,species==spp), xlab="age", ylab="leaf lifespan (leaf births)", cex.axis=.8, las=1,
     main=labels.spp[spp], xlim=c(1,35), ylim=c(0,8), log="x", pch=16)
  points(LL_birth_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8)
  points((LL_birth_mean+LL_birth_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=25)
  points((LL_birth_mean-LL_birth_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=24)
}

@@

```{r, echo=FALSE, message=FALSE}
#plot(LMA_mean ~ as.factor(species), data=subset(SummarySppAge,age>3), xlab="species", ylab="LMA", cex.axis=.8, las=1)
#mod <- lm(LMA ~ age +species + age*species, LMA)
#summary(mod)

@@

plot_LMAvAge <- function(spp) {
  plot(LMA ~ age, data=subset(LMA,species==spp), xlab="age", ylab="LMA", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp])
  points(LMA_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8)
  points((LMA_mean+LMA_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=25)
  points((LMA_mean-LMA_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=24)
  }
  #I want to add a line of code that plots over the means for each age. I did this above by calling in another file, but wondering about on-the-spot calculations

@@

for(n in names(labels.spp)) {
  plot_LMAvAge(n)
  mod_LMAvAge(n)
  summary(mod_LMAvAge(n))
}
```

@@

plot_lfcountvAge <- function(spp) {
  plot(lvs_end_total ~ age, data=subset(SummaryInd,species==spp), xlab="age", ylab="leaf count along shoot", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp])
  points(lvs_end_total_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8)
  points((lvs_end_total_mean+lvs_end_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=25)
  points((lvs_end_total_mean-lvs_end_se) ~ age, data=subset(SummarySppAge, species==spp), cex=1.5, pch=24)
}

mod_lfcountvAge <- function(spp) {
  lm(lvs_end ~ age, data=subset(SummaryInd,age>2.4 & species==spp))
}

for(n in names(labels.spp)) {
  plot_lfcountvAge(n)
  mod_lfcountvAge(n)
  summary(mod_lfcountvAge(n))
}
```
